% ============================================================================
%  Acknowledgments
% ============================================================================

\chapter*{Acknowledgments}
\addcontentsline{toc}{chapter}{Acknowledgments}

The hardest part of writing this report is not the technical content --- it is finding the right words to thank everyone who helped make this project a reality.

First and foremost, I want to express my deepest gratitude to my academic supervisor, \textbf{M.~Slim Abd Elbari}, for his guidance, his valuable advice, and his constant availability throughout this project. His expertise and his constructive feedback helped me stay on the right track, both technically and academically. I truly appreciate the time and effort he invested in reviewing my work.

I would also like to thank \textbf{M.~Hichem Chaaben}, Director of the Institute, for the excellent learning environment he has built at ISITCOM. My thanks also go to \textbf{M.~Slim Abdelbari}, Head of Department, for his support and follow-up throughout our academic journey.

I am deeply grateful to the entire \textbf{teaching staff at ISITCOM Hammam Sousse} for the solid education they provided us over the years. The knowledge and skills I gained from their courses are what made this project possible in the first place.

To the \textbf{members of the jury}: I sincerely thank you for accepting to evaluate this work. Your time, your expertise, and your feedback are deeply valued. Please accept, through this report, my highest respect and appreciation.

A special thank you goes to my \textbf{family} --- my parents, my brothers and sisters, and everyone close to me --- for their unconditional patience, love, and encouragement. They believed in me throughout this entire journey, and I could not have done it without them.

Finally, I want to thank my \textbf{friends and classmates} for the countless discussions, the shared challenges, and the moral support that made this experience not just productive, but truly memorable.

\vspace{1cm}
\begin{flushright}
\textit{Ridha Gnounou}\\
\textit{Hammam Sousse, April 2026}
\end{flushright}
