% ============================================================================
%  Dedication
% ============================================================================

\thispagestyle{empty}
\vspace*{1cm}

\begin{center}
{\Large\itshape Dedication}
\end{center}

\vspace{1cm}

\noindent I dedicate this work to all the people who have been by my side throughout this journey, and without whom none of this would have been possible.

\vspace{0.6cm}

\noindent\textbf{To my mother,}\\
To the woman who gave me everything. Your sacrifices, your patience, and your endless love have been the strongest force behind every step I have taken. If there is any success in this work, it comes from you. You are my first supporter, my safe place, and my biggest motivation. Thank you for always believing in me, even when I doubted myself. This project is yours as much as it is mine.

\vspace{0.6cm}

\noindent\textbf{To my father,}\\
To the man who taught me the value of hard work, honesty, and perseverance. Your wisdom and your quiet confidence in me have been my guide through every challenge. You showed me that nothing worth having comes easy, and that lesson has carried me through this entire project. I hope this work makes you proud.

\vspace{0.6cm}

\noindent\textbf{To my brothers and sisters,}\\
Thank you for the laughs, the encouragement, and for always being there. Your support means more than words can say. I believe in each of you, and I know great things are ahead for all of us.

\vspace{0.6cm}

\noindent\textbf{To my family,}\\
To my grandparents, uncles, aunts, and cousins --- thank you for the love and the warm gatherings that kept me going. This work carries a piece of all of you in it.

\vspace{0.6cm}

\noindent\textbf{To my friends,}\\
To my closest friends who stood by me during the long nights and the stressful days. Your presence turned the hardest moments into bearable ones. Thank you for the motivation, the coffee breaks, and for never letting me give up.

\vspace{0.6cm}

\noindent\textbf{To my classmates and colleagues,}\\
Working alongside you has been one of the best parts of this journey. Thank you for the teamwork, the shared knowledge, and the energy that made every challenge more enjoyable.

\vspace{0.8cm}

\noindent To all of you, from the bottom of my heart --- thank you. This project is the result of your love, your trust, and your constant support.
